Similar Article Detection Using Locality Sensitive Hashing Technique on Wikipedia\titlenote{This effort is part of the CMSC828G, Data Intensive Computing with MapReduce, class project.}

### 2013-05-09 article/sigproc-sp.tex
Added some clarifying text to the ed/jaccard analysis section.
Updates to the report

\author{
\alignauthor Samet Ayhan\\
       \affaddr{University of Maryland, Dept. of Computer Science}\\
       \affaddr{College Park, Maryland 20740}\\
       \email{sayhan@cs.umd.edu}
\alignauthor Joshua Bradley\\
       \affaddr{University of Maryland, Dept. of Computer Science}\\
       \affaddr{College Park, Maryland 20740}\\
       \email{jgbrad1@cs.umd.edu}
\alignauthor Sarah Weissman\\
       \affaddr{University of Maryland, College of Information Studies}\\
       \affaddr{College Park, Maryland 20740}\\
       \email{sew@umd.edu}
}

\date{19 April 2013}

@@

In this paper, we describe a novel similarity detection algorithm that utilizes a locality sensitive hashing (LSH) technique on the MapReduce framework. The algorithm has been designed and implemented to detect similar articles using large Wikipedia dumps, in compressed or uncompressed forms. Experimental results appear to support that our method is able to efficiently and effectively detect similar articles across Wikipedia.
\end{abstract}

@@

Furthermore, Wikipedia has a history of generating controversy over editorial quality and factual correctness. Although some studies have found Wikipedia’s accuracy to rival that of traditional encyclopedias, other studies have found numerous factual errors. Since Wikipedia may be edited anonymously, information is often freely copied between web sources and even between Wikipedia articles. Additionally, many articles in Wikipedia lack citations. There is little mechanism for detecting and correcting such errors. One potential source of error on Wikipedia is factual information that varies over time and is not updated. Wilkinson found that the distribution of article edits has a long tail, meaning that a small number of articles have many edits while most articles have few, and that the number of edits is directly related to article quality. Articles with low edits and low editorial attention, are less likely to be updated when new information is available \cite{wilkinson:wiki}.

@@

In order to come up with a novel algorithm that would allow us to detect near duplicate Wikipedia articles, we decided that we needed to answer the following questions, first:
\begin{itemize}
\item What is the most appropriate LSH technique to apply? Edit distance seems like a reasonable metric for near duplicate sentences, however LSH does not map naturally to the edit distance metric. Currently, no LSH family for edit distance is known to exist.
\item Compare different sentence similarity techniques for effectiveness. 
\item What is the appropriate granularity for analysis? Is comparing multiple consecutive sentences more effective than comparing single sentences? Similarly, what is the best shingle size?
\item How well do techniques for detecting near duplicate documents work when applied at the sentence level?
\item Do near duplicate sentences correspond to factual errors? Can these errors be automatically detected and/or corrections suggested?
\end{itemize}

@@

The rest of this paper is organized as follows: In Section 2, we present related work, in Section 3, we explain the algorithm utilizing locality sensitive hashing, in Section 4, we present the working application within MapReduce framework. In Section 5, we discuss our experiments with various Wikipedia datasets. The final section contains concluding remarks and future work.

@@

\section{Algorithm Utilizing Locality Sensitive Hashing}

@@

Describe MapReduce.

@@

Our implementation is built on top of the Apache Hadoop MapReduce framework, using utilities from the Cloud9 library (\url{https://github.com/lintool/Cloud9}) for parsing Wikipedia page text from the Wikipedia XML dump format. Sentences are parsed out over each document using a regular expression. Sentences that are too large or too small are discarded. The family of hash functions is implemented using a ``Multiply Shift''  \cite{} hashing scheme and is generated from a random seed using the Java Random class. The hash family is also parametrized by hash output key size and this value can be configured to affect the size of the output signatures.

@@

In order to estimate Jaccard similarity of sentences $X$ and $Y$ in terms of edit distance we assume that $X$ and $Y$ are both of length $N$ and that their shingle sets are of maximum size (i.e. all shingles are unique). Also assume that $X$ and $Y$ have edit distance $e$.  Additionally, to simplify our calculations, we assume that edits are localized so that no shingle overlaps more than one edit.

@@

All of our experiments were run against an XML dump of current English-language Wikipedia (enwiki) articles (as of 4/2013) obtained from dumps.wikimedia.org. The enwiki dump is approximately 40G in size and contains 4,113,785 articles (after excluding certain non-article page types). Running the first four steps of our application (up to but not including the final edit distance filtering pass) against the enwiki dump on a shared cluster of 16 nodes with (15 GB RAM and 4 cores each) took XXXX (HOW MUCH TIME DID THIS TAKE, SAMET?) minutes.

@@

Cluster sizes ranged from 2 to 28,008. (See Figure \ref{clust} for a breakdown of counts by cluster size.) Most sentence clusters are small, but the distribution of cluster sizes exhibits a long tail. As a result, while 99\% of our clusters have size below or equal to 30, about 1/3 of the output sentence/article pairs fall into clusters with size greater than 30. One explanation for this distribution of sentences is the phenomenon we call ''templatification,'' which we describe in more detail below.)

@@

 \emph{Templatification} occurs when one sentence is an exact duplicate of another sentence, but the main topic has been changed by only a few pieces of key information having been modified. For example, it could be said that the following two sentences follow a particular \emph{template}.
\begin{itemize}[noitemsep,nolistsep]
\item Bush had an approval rating of 22\% by the end of his term in 2008.
\item Obama had an approval rating of 56\% by the end of his term in 2012.
\end{itemize}
From the perspective of editing/writing Wikipedia articles, this is a strong indicator that the author copied the original sentence and modified the relevant information to pertain to a different topic. What remains shared between both sentences though is the inherent structure.

\emph{Direct Copying} occurs when the same sentence has been copied into another (or possibly the same) article. This can happen for many reasons, including articles that deal with similar subjects, or articles that are subtopics of other topics. (SOME EXAMPLES?). Similarly, \emph{Direct Copying with Copy Editing} occurs when the same sentence has been copied into another article and then one or both sentences has been edited for style or clarity. For example:

\begin{itemize}[noitemsep,nolistsep]
\item EXAMPLE HERE.
\item EXAMPLE HERE.
\end{itemize}

@@

\begin{itemize}[noitemsep,nolistsep]
\item EXAMPLE HERE.
\item EXAMPLE HERE.
\end{itemize}

@@

By manually counting a small sample of NNNN output clusters, we came up with the following breakdown of sentence types:

INSERT CLASSIFY CLUSTER BREAKDOWN HERE

@@

JOSH, CAN YOU LOOK UP WHAT THE ACTUAL ARTICLES ARE? THAT MIGHT BE INTERESTING.

\section{Conclusion and Future Work}
This paper presents our work on near-duplicate detection of Wikipedia articles at various granularity levels using LSH technique. Our novel MapReduce algorithm tackles the problem by transforming it into more manageable pieces where each piece is handled in parallel. Our unique contribution includes empirical analysis of signatures belonging to Wikipedia articles at the sentence level. Experimental results verify that the proposed method is able to effectively and efficiently detect similar articles.

@@

\section{Acknowledgments}
We would like to thank all of the anonymous contributors. We would especially like to thank Jimmy Lin for his advise and directions.

>>>>

\author{
\alignauthor Samet Ayhan\\
       \affaddr{University of Maryland}\\
       \affaddr{Dept. of Computer Science}
       \email{sayhan@cs.umd.edu}
\alignauthor Joshua Bradley\\
       \affaddr{University of Maryland}\\
       \affaddr{Dept. of Computer Science}
       \email{jgbrad1@cs.umd.edu}
\alignauthor Sarah Weissman\\
       \affaddr{University of Maryland}\\
       \affaddr{College of Information Studies}\\
       \email{sew@umd.edu}
}

\date{9 May 2013}

@@

In this paper, we describe a novel similarity detection algorithm that utilizes a locality sensitive hashing (LSH) technique on the MapReduce framework. The algorithm has been designed and implemented to detect similar articles using large Wikipedia dumps. Experimental results appear to support that our method is able to efficiently and effectively detect similar articles across Wikipedia.
\end{abstract}

@@

Furthermore, Wikipedia has a history of generating controversy over editorial quality and factual correctness. Although some studies have found Wikipedia's accuracy to rival that of traditional encyclopedias, other studies have found numerous factual errors. Since Wikipedia may be edited anonymously, information is often freely copied between web sources and even between Wikipedia articles. Additionally, many articles in Wikipedia lack citations. There is little mechanism for detecting and correcting such errors. One potential source of error on Wikipedia is factual information that varies over time and is not updated. Wilkinson found that the distribution of article edits has a long tail, meaning that a small number of articles have many edits while most articles have few, and that the number of edits is directly related to article quality. Articles with low edits and low editorial attention, are less likely to be updated when new information is available \cite{wilkinson:wiki}.

@@

In order to come up with a novel algorithm that would allow us to detect near-duplicate sentences within Wikipedia articles, we decided that we needed to answer the following questions, first:
\begin{itemize}[noitemsep,nolistsep]
\item What is the most appropriate LSH technique to apply? Edit distance seems like a reasonable metric for near duplicate sentences, however LSH does not map naturally to the edit distance metric.
\item Compare different sentence similarity techniques for effectiveness. 
\item What is the appropriate granularity for analysis? Is comparing multiple consecutive sentences more effective than comparing single sentences? Similarly, what is the best shingle size?
\item How well do techniques for detecting near duplicate documents work when applied at the sentence level?
\item Do near duplicate sentences correspond to factual errors? Can these errors be automatically detected and/or corrections suggested?
\end{itemize}

@@

The rest of this paper is organized as follows: In Section 2, we present related work, in Section 3, we explain the algorithm utilizing locality sensitive hashing, in Section 4, we discuss parameter tuning for Minhash. In Section 5, we present the working application within the MapReduce framework. In Section 6, we discuss our experiments with various Wikipedia datasets. The final section contains concluding remarks and future work.

@@

\section{Algorithm Utilizing Locality \\ Sensitive Hashing}

@@

DESCRIBE MAPREDUCE.

@@

Our implementation is built on top of the Apache Hadoop MapReduce framework, using utilities from the Cloud9 library (\url{https://github.com/lintool/Cloud9}) for parsing Wikipedia page text from the Wikipedia XML dump format. Sentences are parsed out over each document using a regular expression. Sentences that are too large or too small are discarded. The family of hash functions is implemented using a ``Multiply Shift''  CORRECT CITATION GOES HERE hashing scheme and is generated from a random seed using the Java Random class. The hash family is also parametrized by hash output key size and this value can be configured to affect the size of the output signatures.

@@

In order to estimate Jaccard similarity of sentences $X$ and $Y$ in terms of edit distance we assume that $X$ and $Y$ are both of length $N$ and that their shingle sets are of maximum size (i.e. all shingles are unique). Also assume that $X$ and $Y$ have edit distance $e$.  We limit type of edits to changing characters in place. Additionally, to simplify our calculations, we assume that edits are localized so that no shingle overlaps more than one edit.

@@

All of our experiments were run against an XML dump of current English-language Wikipedia (enwiki) articles (as of 4/2013) obtained from dumps.wikimedia.org. The enwiki dump is approximately 40G in size and contains 4,113,785 articles (after excluding certain non-article page types). Running the first four steps of our application (up to but not including the final edit distance filtering pass) against the enwiki dump on a shared cluster of 16 nodes with (15 GB RAM and 4 cores each) took $\approx 2$ hours.

@@

Cluster sizes ranged from 2 to 28,008. (See Figure \ref{clust} for a breakdown of counts by cluster size.) Most sentence clusters are small, but the distribution of cluster sizes exhibits a long tail. As a result, while 99\% of our clusters have size below or equal to 30, about 1/3 of the output sentence/article pairs fall into clusters with size greater than 30. One explanation for this distribution of sentences is the phenomenon we call ''templatification,'' which we describe in more detail below.

@@

 \emph{Templatification} occurs when one sentence is an exact duplicate of another sentence, but the main topic has been changed by only a few pieces of key information having been modified. For example, it could be said that the following two sentences follow a particular \emph{template}.
\begin{itemize}[noitemsep,nolistsep]
\item Bush had an approval rating of 22\% by the end of his term in 2008.
\item Obama had an approval rating of 56\% by the end of his term in 2012.

\end{itemize}
From the perspective of editing/writing Wikipedia articles, this is a strong indicator that the author copied the original sentence and modified the relevant information to pertain to a different topic. What remains shared between both sentences though is the inherent structure.

\emph{Direct Copying} occurs when the same sentence has been copied into another (or possibly the same) article. This can happen for many reasons, including articles that deal with similar subjects, or articles that are subtopics of other topics. (SOME EXAMPLES?). Similarly, \emph{Direct Copying with Copy Editing} occurs when the same sentence has been copied into another article and then one or both sentences has been edited for style or clarity. For example:
\begin{itemize}[noitemsep,nolistsep]
\item Bush had an approval rating of 22\% by the end of his term in 2008.
\item Bush, a US President, had an approval rating of 22\% by the end of his term in 2008.
\end{itemize}

@@

\begin{itemize}[noitemsep,nolistsep]
\item Neotropical Ichthyology 11 (1): 73-80.
\item Neotropical Ichthyology 10 (2): 245-253.
\end{itemize}

@@

By manually counting a small sample of 824 output clusters, we came up with a breakdown of sentence types in Table 1.

\begin{table}
\centering
\caption{Sentence Breakdown Counts}
\begin{tabular}{| l | c |}
\hline
Not Similar & 3 \\ \hline
Factual Drift & 65 \\ \hline
Template & 52 \\ \hline
Reference & 23 \\ \hline
Copy Edit & 214 \\ \hline
Other & 12 \\ \hline
Identical & 455 \\
\hline
\end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Top 10 Article Pair Titles}
\begin{tabular}{| l | c |}
    \hline
    \textbf{Article ID} & \textbf{Title} \\ \hline
    21986867 & 2000 New Year Honours \\ \hline
    21986897 & 1999 New Year Honours \\ \hline
    3951586 & Instruments of the United Kingdom, 1994 \\ \hline
    3951664 & Instruments of the United Kingdom, 1995 \\ \hline
    1903966 & Management of baldness \\ \hline
    38088660 & Management of androgenic alopecia \\ \hline
    1503105 & Jayne Torvill \\ \hline
    1503115 & Christopher Dean \\ \hline
    184843 & Henry VI, Part 1 \\ \hline
    184846 & Henry VI, Part 2 \\ \hline
    184847 & Henry VI, Part 3 \\ \hline
    10701462 & List of birds of Zambia \\ \hline
    10930255 & List of birds of Zimbabwe \\ \hline
    10930111 & List of birds of Mozambique \\ \hline
    48711 & Oregon Trail \\ \hline
    31518964 & History of the Oregon Trail \\
    \hline
\end{tabular}
\end{table}

\section{Conclusion and Future Work}
This paper presents our work on near-duplicate detection of Wikipedia articles at the sentence-granularity level using MinHash, a LSH technique. Our novel MapReduce algorithm tackles the problem by transforming it into more manageable pieces where each piece is handled in parallel. Our unique contribution includes empirical analysis of signatures belonging to Wikipedia articles at the sentence level. Experimental results verify that the proposed method is able to effectively and efficiently detect similar articles.

@@

\section{Acknowledgments}
We would especially like to thank Dr. Jimmy Lin for his advise and directions.
